\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-06-27 write_up/mixfishsim_ecolmod.tex
Text updates

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions
using species non-selective fishing gear. \added[id=PD]{In doing so} fisheries
\deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota which may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{This discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}}\citep{Kennelly2002, Catchpole2008,
	Bellido2011}.  \\

\replaced[id=PD]{Changes to spatial fishing patterns have}{Use of spatial
	management as a tools} been proposed as a method to reduce discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.  \deleted[id=PD]{its}
implementation of avoidance measures is, however, restricted by lack of
knowledge of fish and fishery spatiotemporal dynamics and understanding of the
scale at which processes become important for management. Understanding the
correct scale for spatial measures is crucial for implementation at a
resolution that ensures effective management \citep{Dunn2016} while minimising
economic impact.  For example, a scale that promotes species avoidance for
vulnerable or low quota species while allowing continuance of sustainable
fisheries for available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} an
appropriate scale has been a challenge in the past that has led to ineffectual
measures with unintended consequences such as limited impact towards the
management objective or increased benthic impact on previously unexploited
areas (e.g. the cod closure in the North Sea
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatial information has \added[id=PD]{since} become available through
the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}.  Such
information is, however, derived from an inherently biased sampling programme,
targeted fishing. \deleted[id=PD]{ Further, fishers generally only recorded
	landings (not catch) on a daily basis\todo{\added[id=JJ]{Its not clear
			what the problem is: landings collected on daily basis
			or landings recorded rather than catch}}.  This leads
	to questions about the validity of inference that can be drawn from
	landings data assigned to VMS activity pings.}\todo{\added[id=JJ]{This
		comes as a surprise: I thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

@@

We i) develop a simulation model where population dynamics are highly-resolved
in space and time to answer these questions\replaced[id=PD]{. Being}{ and are}
known \added[id=PD]{directly} rather than inferred from sampling or commercial
catch\added[id=PD]{, we can use the population model to evaluate how inference
	from fisheries-dependent and fisheries independent sampling relates to
	the real population structure}. We ii) compare at different spatial and
temporal aggregations of the simulated population distributions to samples from
fisheries-dependent and fisheries independent catches. We then iii) simulate a
fishery closure to protect a species based on different spatial and temporal
data aggregations. \\

We compare the catch composition from fisheries independent and dependent
sampling against the real population composition, as well as the theoretical
"benefit" to the population spatial closures using the different data sources.
We use these evaluations to draw inference on the utility of commercial data in
supporting management decisions. \todo{\added[id=JJ]{If the paper has two goals
		this should be clear from the start, but may be better over 2
		MSs}\added[id=PD]{I would like to keep both parts, but have
		made clearer in how its set out.  The closure scenarios form
		validation of the data aggregation, rather than effectiveness
		of the closures themselves - so its a continuation of the same
		question in my eyes}}\deleted[id=PD]{.  Further, we extend our
	analysis to a range of spatial and temporal scales to assess the impact
	of these processes on the success of the management measure.}\\

@@

\replaced[id=PD]{A}{We developed and implemented a simulation model with a}
modular event-based \replaced{simulation model was developed with}{approach,
	where} \added[id=PD]{sub-}modules \deleted[id=PD]{are} implemented on
independent time-scales appropriate to capture the characteristic of the
\replaced[id=PD]{different processes}{process modelled} (Figure 1).
\added[id=PD]{The following sub-modules were included to capture the full
	system: 1) Population dynamics, 2) Recruitment dynamics, 3) Population
	movement, 4) fishery dynamics.}\\

@@

\replaced[id=PD]{In our model system p}{P}opulation movement is driven by
random (diffusive) and directed (advective) processes and we incorporate
characterisation of a number of different fishing \added[id=PD]{fleet dynamics}
exploiting four fish populations with different spatial and population
demographics. We use the model to simulate \replaced[id=PD]{50}{40} years of
exploitation of the fish populations.

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen to discretise continuous population processes
	on a biologically relevant and computationally tractable timescale.}
\replaced[id=PD]{Under the population dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and \deleted[id=PD]{fish}
recruited \added[id=PD]{fish} \deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment. Biomass for each cell $c$ is incremented each day $d$ as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}):
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be approximately equal to
$e^{-K}$ when $K$ is the growth rate from a von Bertalanffy logistic growth
model \citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In this study, the
Beverton-Holt form of stock recruit relationship was used for all populations.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations on a 2d
grid\replaced[id=JJ]{. An}{, with an} advection-diffusion process
\deleted[id=JJ]{to} control\added[id=JJ]{led} \deleted[id=JJ]{how the}
population\deleted[id=JJ]{s} move\replaced[id=JJ]{ment}{d}, with a time-varying
temperature covariate used to change the spatial bounds of suitable habitat
on a weekly time-step\todo{\added[id=JJ]{What have a temperature covariate? Could just
		use time}\added[id=PD]{Was intended as some biological meaning
		- species thermal tolerances load onto the temperature effect -
		so could be different per species}}.  \deleted[id=JJ]{This was
	intended to balance realism in population movement, capturing the main
	directed and random processes, and practicality of modelling the
	population rather than individual fish.} \\

For \deleted[id=PD]{the}habitat we first define\added[id=PD]{d} a Gaussian
random field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian. The distribution is
specified by its \textit{mean function}, $\mu(c) = E[S(c)]$ and its
\textit{covariance function}\todo{\added[id=JM]{Introduce the gamma function,
		and why this covariance structure? Why correlate values in the
		random field?}\added[id=PD]{to allow populations to have
		different aggregation densities: have tried to clarify}},
$\gamma(c,c') = Cov\{S(c),S(c')\}$ \citep{Diggle2007}.\\

@@

The temperature field \replaced[id=PD]{was}{is} defined on a gradient from a
South-Westerly to North-Easterly direction, with temperature in each cell
changing gradually on a week-by-week basis so that initially high temperature
areas cycle\added[id=PD]{d} to lower temperatures and low temperature areas
\textit{vice versa}. Each population $p$ \replaced[id=PD]{wa}{i}s assigned a thermal
tolerance with mean, $\mu\added[id=PD]{_{p}}$ and variance,
$\sigma^2\added[id=PD]{_{p}}$ so that each cell and population temperature
suitability is defined that:
\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ for
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final process resulted in independent populations structure
	and movement patterns, with population movement occurring on a weekly
	basis. This process approximated the demographic shifts in fish
	populations throughout a year with seasonal spawning patterns (e.g.
	Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these element implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution (\cite{Bailey2018}).

@@

Fishing locations within a trip are \added[id=PD]{initially} determined by a
modified random walk process. \added[id=PD]{As the simulation progresses the
	within-trip decision become gradually more influenced by experience
	gained from past fishing locations (as per the initial trip-level
	location choice), moving location choice towards areas of higher
	perceived profit.} A random walk was chosen for the exploratory fishing
process as it is the simplest assumption commonly used in ecology to describe
\added[id=PD]{optimal} animal\deleted[id=PD]{movement which}
search\replaced[id=PD]{ strategy}{ing} for \added[id=PD]{exploiting}
homogeneously distributed prey about which there is uncertain knowledge
\citep{Viswanathan1999}. In a random walk, movement is a stochastic process
through a series of steps\added[id=JJ]{. These steps have a length, and a
	direction} that can either be equal in length or take some other
functional form. The direction of the random walk was also correlated (known as
`persistence') providing some overall \deleted[id=PD]{location of}directional
movement \citep{Codling2008} \deleted[id=PD]{or uncorrelated}. \\

>>>>

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions
and using species non-selective fishing gear. \added[id=PD]{In doing so}
fisheries \deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011}.  \\

\replaced[id=PD]{Changes to spatial fishing patterns have}{Use of spatial
	management as a tools} been proposed as a method to reduce discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.  \deleted[id=PD]{its}
Implementation of avoidance measures is, however, restricted by lack of
knowledge of fish and fishery spatiotemporal dynamics and understanding of the
scale at which processes become important for management. Understanding the
correct scale for spatial measures is crucial for implementation at a
resolution that ensures effective management \citep{Dunn2016} while minimising
economic impact.  For example, a scale that promotes species avoidance for
vulnerable or low quota species while allowing continuance of sustainable
fisheries for available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} an
appropriate scale has been a challenge in the past that has led to ineffectual
measures with unintended consequences such as limited impact towards the
management objective or increased benthic impact on previously unexploited
areas (e.g. the cod closure in the North Sea
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatial information has \added[id=PD]{since} become available through
the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}.  Such
information is, however, derived from an inherently biased sampling programme,
targeted fishing, where fishers establish favoured fishing grounds first
through an explore-exploit strategy \cite{Bailey2018} and then use experience
to return to areas of high catch. \deleted[id=PD]{ Further, fishers generally
	only recorded landings (not catch) on a daily
	basis\todo{\added[id=JJ]{Its not clear what the problem is: landings
			collected on daily basis or landings recorded rather
			than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time \replaced[id=PD]{. Being}{ and
	are} known \added[id=PD]{directly} rather than inferred from sampling
or commercial catch\added[id=PD]{, we can use the population model to evaluate
	how inference from fisheries-dependent and fisheries independent
	sampling relates to the real population structure}. We ii) compare, at
different spatial and temporal aggregations, the simulated population
distributions to samples from fisheries-dependent and fisheries independent
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. We use these evaluations
to draw inference on the utility of commercial data in supporting management
decisions. \todo{\added[id=JJ]{If the paper has two goals this should be clear
		from the start, but may be better over 2 MSs}\added[id=PD]{I
		would like to keep both parts, but have made clearer in how its
		set out.  The closure scenarios form validation of the data
		aggregation, rather than effectiveness of the closures
		themselves - so its a continuation of the same question in my
		eyes}}\deleted[id=PD]{.  Further, we extend our analysis to a
	range of spatial and temporal scales to assess the impact of these
	processes on the success of the management measure.}\\

[We find..]

@@

\replaced[id=PD]{A}{We developed and implemented a simulation model with a}
\replaced{simulation model that is modular and discrete-event based was
	developed. This approach enables efficient computation by allowing
	for}{approach, where} \added[id=PD]{sub-}modules \deleted[id=PD]{are}
implemented on time-scales appropriate to capture the characteristic of the
\replaced[id=PD]{different processes}{process modelled} (Figure 1).
\added[id=PD]{The following sub-modules were included to capture the full
	system: 1) Population dynamics, 2) Recruitment dynamics, 3) Population
	movement, 4) fishery dynamics.}\\

@@

\replaced[id=PD]{In the model system p}{P}opulation movement is driven by
random (diffusive) and directed (advective) processes and we incorporate
characterisation of a number of different fishing \added[id=PD]{fleet dynamics}
exploiting four fish populations with different spatial and population
demographics. The following describes the implementation of each of the
sub-modules.

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen to discretise continuous population processes
	on a biologically relevant and computationally tractable timescale.}
\replaced[id=PD]{Under the population dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and\deleted[id=PD]{fish}
recruited \added[id=PD]{fish} \deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment. Biomass for each cell $c$ is incremented each day $d$ as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}):
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be equal to $e^{-K}$ when $K$ is
the growth rate from a von Bertalanffy logistic growth model
\citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In our example
application the Beverton-Holt form of stock recruit relationship was used for
all populations though either functional form can be chosen.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations on a 2d grid. \\

\deleted[id=PD]{the habitat}We first define\added[id=PD]{d} a Gaussian random
field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian. The distribution is
specified by its \textit{mean function}, $\mu(c) = E[S(c)]$ and its
\textit{covariance function}\todo{\added[id=JM]{Introduce the gamma function,
		and why this covariance structure? Why correlate values in the
		random field?}\added[id=PD]{to allow populations to have
		different aggregation densities: have tried to clarify}},
$\gamma(c,c') = Cov\{S(c),S(c')\}$ \citep{Diggle2007}.\\

@@

\replaced[id=JJ]{An}{, with an} advection-diffusion process \deleted[id=JJ]{to}
control\added[id=JJ]{led} \deleted[id=JJ]{how the} population\deleted[id=JJ]{s}
move\replaced[id=JJ]{ment}{d}, with a time-varying temperature covariate used
to change the interaction between time and suitable habitat on a weekly
time-step\todo{\added[id=JJ]{What have a temperature covariate? Could just use
		time}\added[id=PD]{Was intended as some biological meaning -
		species thermal tolerances load onto the temperature effect -
		so could be different per species}}.  \deleted[id=JJ]{This was
	intended to balance realism in population movement, capturing the main
	directed and random processes, and practicality of modelling the
	population rather than individual fish.} Each population $p$
\replaced[id=PD]{wa}{i}s assigned a thermal tolerance with mean,
$\mu\added[id=PD]{_{p}}$ and variance, $\sigma^2\added[id=PD]{_{p}}$ so that
each cell and population temperature suitability is defined that:
\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ for
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final process resulted in population structures and movement
	patterns unique to each species, with population movement occurring on
	a weekly basis. The decision to model population movement on a weekly
	timescale was to reflect that fish tend to aggregate in species
	specific locations and range within a week is fairly limited [REF!!].
	Therefore this process approximated the demographic shifts in fish
	populations throughout a year with seasonal spawning patterns (e.g.
	Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these element implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution (\cite{Bailey2018}). The decision to use an individual
based model for fishing vessels was taken because fishers as a group tend to
show heterogeneity and individual rather than group dynamics. Thus this was the
fleet dynamics is the productive of individual experiences rather than
pre-ordained group dynamics.

@@

Fishing locations within a trip are \added[id=PD]{initially} determined by a
modified random walk process. \added[id=PD]{As the simulation progresses the
	within-trip decision become gradually more influenced by experience
	gained from past fishing locations (as per the initial trip-level
	location choice), moving location choice towards areas of higher
	perceived profit.} A random walk was chosen for the exploratory fishing
process as it is the simplest assumption commonly used in ecology to describe
\added[id=PD]{optimal} animal\deleted[id=PD]{movement which}
search\replaced[id=PD]{ strategy}{ing} for \added[id=PD]{exploiting}
homogeneously distributed prey about which there is uncertain knowledge
\citep{Viswanathan1999}. In a random walk, movement is a stochastic process
through a series of steps\added[id=JJ]{. These steps have a length, and a
	direction} that can either be equal in length or take some other
functional form. The direction of the random walk was also correlated (known as
`persistence') providing some overall \deleted[id=PD]{location of}directional
movement \citep{Codling2008}\deleted[id=PD]{or uncorrelated}. \\
